% Plantilla de memoria Universidad de los Andes
% para usuarios de pdfLaTeX.
%
% Jaime Cisternas, Santiago, junio de 2004, agosto 2010
%
% Necesita todos los archivos *.tex llamados en las
% sentencias \input{} (cap1, cap2,...)
% ademas de buc.bib y las figuras.
%
% El archivo setspace.sty permite especificar el espaciado
% entre las lineas. Los archivos newapa.sty y newapa.bst
% especifican el formato de la bibliografia.
%
% Para compilar :                    pdflatex memoria
% Para generar la bibliografia :     bibtex memoria, pdflatex memoria.tex (dos veces)

\documentclass[12pt]{report}

% El archivo siguiente contiene todos los detalles
% finos del formato de la memoria.
% No es necesario modificarlo.


\input{core/preambulo.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% LAS SIGUIENTES LINEAS DEBEN SER MODIFICADAS POR EL AUTOR DE LA MEMORIA
% SI EL TITULO ES MUY LARGO PUEDE SER CORTADO CON \linebreak

\newcommand{\nombreautor}{Luca Appiani Caro}
\newcommand{\mes}{\monthname}
\newcommand{\anio}{\the\year}
\newcommand{\titulo}{Formulation of an asset pricing model for cryptocurrencies}
\newcommand{\nombreprofuno}{Javier Mella}
\newcommand{\nombreprofdos}{Name of the Dean's Representative}
\newcommand{\nombreproftres}{Name of the Visiting Professor}
\newcommand{\codigo}{ING-IN-001/11}

% LAS SIGUIENTES DEFINICIONES TAMBIEN SON NECESARIAS.
% ES FUNDAMENTAL QUE EL TEXTO ESTE ENCERRADO ENTRE
% PARENTESIS DE LLAVE.

\newcommand{\resumen}{
Cryptocurrency has gained significant popularity in recent years due to various factors, including attractive returns and new investment opportunities. Additionally, the technologies underlying these assets are highly interesting to investors worldwide. Therefore, research aimed at identifying the factors driving cryptocurrency returns can be highly beneficial for both individual investors and financial institutions, which currently hold substantial investments in this type of assets.

The main objective of this dissertation is to create and confirm a Portfolio Markowitz-inspired model to evaluate cryptocurrency returns in order to improve comprehension and assist in making informed investment choices.

The methodology is comprised of five main steps. First, the mathematical formulation of the model was done, which was inspired by the research of \parencite{luo2017social}. The procedure is similar, although there were some steps that were not detailed in the original paper, but in this case the interpretation of various variables is changed, in order to fit the cryptocurrency focus. Then, a dataset of returns from cryptocurrencies was built, and to ensure the validity of said returns, some statistical tests were conducted. 

Following this, statistical tests were conducted, including Fama-MacBeth regressions and the Gibbons, Ross, and Shanken (GRS) test, to evaluate the model's capacity and validity in explaining the cross-section of cryptocurrency returns. Subsequently, conclusions and recommendations were drawn from these tests.

The most important results include the model passing the GRS test at a 5\% significance level, but not at 1\%. The Fama-MacBeth regressions yielded poor results, with factors not statistically significant and an average R-squared of 33\%. Principal conclusions indicate that while theory-based models may explain asset returns, empirical models with more factors outperform them when tested with real data.
}

\newcommand{\agradecimientos}{
% Muchas gracias a todas aquellas personas que quizás
% sin saberlo y sin mala voluntad de su parte
% colaboraron a la finalización de esta investigación... 

%En esta página se puede nombrar a toda clase de personas,
%amigos y parientes, sin olvidar al profesor guía y
%a la secretaria de la Facultad.
}

\newcommand{\dedicatoria}{\it
I dedicate this work to ...
% mis papás, a mis abuelitos,
% a mi nana y a mi tía del Kinder.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Por favor coloque todas las definiciones de simbolos a continuacion
% y no en los archivos de capitulos. Esto evitara la  existencia de
% multiples definiciones para las mismas palabras.

\newcommand{\mydef}
	{\stackrel{\mathrm{def}}{=}}
\newcommand{\e}
	{\hbox{\large{e}}}
\newcommand{\mi}
	{\hbox{\large{i}}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\refname}{References}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\label{start}

% Genera las paginas de titulo, copyright, resumen,
% agradecimientos, dedicatoria, indices,...

% No es necesario modificar

\input{core/primeras_paginas.tex}
%\listoftodos
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{chapters/chapter01.tex}  
\cleardoublepage

\input{chapters/chapter02.tex}
\cleardoublepage

\input{chapters/chapter03.tex}
\cleardoublepage

\input{chapters/chapter04.tex}
\cleardoublepage

\input{chapters/chapter05.tex}
\cleardoublepage

\input{chapters/chapter06.tex}
\cleardoublepage


% Bibliography
\printbibliography[heading=bibintoc, title = {References}]



\appendix % seguido de archivos de apendices

\cleardoublepage

% Escribe 'Anexos' en el Indice General
\addcontentsline{toc}{chapter}{Appendix}

\input{attachements/anexo_a}    % archivo de apendice

%\input{apendice_b}    % otro archivo de apendice

\label{end}
\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
